|

### 2021-08-05 appendix/appendix.Rmd
adding cost to govern

\subsection{A Model of Legislature Size and Service Provision}

@@

We solve the game for a Sub-game Perfect Nash Equilibrium using backward induction. Although the game tree may extend infinitely, a Stationary Sub-game Perfect Equilibrium can be reached with no delay in the policy implementation, and the mayoral proposal being successful.

\begin{definition}
The solution for the game consists in a pair of XXXX.
\end{definition}

Solving by backward induction, suppose that there were $k$ prior rejections, and the municipal resources shrank by $\delta^{k}R$. As at this stage the politicians are similar with regard to political affiliation, we can drop the party label.\footnote{In the extensions, we solve the game assuming that the distribution remained partisan even for the councilors. The results are similar, with the only difference that when partisanship is relevant for councilors, the expected provision of public goods is higher than zero.} A given councilor is recognized as the proposer with $1/N$ probability. The recognized councilor has to convince at least one-half of the council members to agree with her policy proposal. Therefore, a legislator $i$, that receives an offer from a proposer ($x_i, t_i$), at round $k$, will accept the offer if it is at least as good as waiting for the next round.

@@

We now have the entire strategy profile for the city councilors, provided that they reject the mayor's proposal. Proceeding by backward induction, we solve for the mayor's proposal that wins the support of the majority of the council. In the mayor's proposal policy concerns come into play and the offer for a government an aligned councilor is different than the offer for an opposition councilor. For a government aligned councilor, assuming that the councilor is the pivotal players, the accepted offer $(\underline{x}_m^g, \underline{t}_m^g)$ has to be better than rejecting the proposal and waiting for the next round:

@@

Note that $\underline{x}(1)$ is the offer of private goods if the councilor rejects the mayors offer, so that $k = 1$. As the benefit of retention $B$ is higher than the cost of providing the targeted goods $\underline{t}$, the optimal provision of targeted goods satisfy the threshold for reelection: $t^* = t_m^g = t_i = \underline{t}$. Rewriting the equations with this info leads to:

@@

Solving for $x_m^g$ leads us to:

\[
x_m^g \ \geq \ -p +\dfrac{1}{2N+1}\left[2\delta R \left(1 + \dfrac{1-\delta}{2N}\right)-N(\underline{t} + B) + B\right] \equiv \underline{x}_m^g
\]

For the opposition politician, the offer is equal to:

,

.

>>>>

# A Model of Legislature Size and Service Provision

## Full model

@@

This is a bargaining game, along the lines of @rubinsteinXXXX. For political science models similar to this model, see @baronfereXXXX and @cliffsXXXX.

The solution for this game consists in finding a Stationary Sub-game Perfect Nash Equilibrium using backward induction. Although the game tree may extend infinitely, a Stationary Sub-game Perfect Equilibrium assumes that if a politician accepts an offer at time $t+1$, she should accept the offer at time $t$. This gives us an objective of finding the offer that will be accepted at any point in the game. Moreover, applying this reasoning we can find the optimal offer at $t=0$, which represents no delay in policy implementation. The mayor calibrates the offer for approval offering the benefits that councilors would receive, should they reject the offer and enter the reversal stage.

\begin{definition}
The solution for the game consists in a pair of strategic policy vectors for the mayor, and the city councilors. The mayor's policy vector is comprised of a public goods provision $g^*$ and XXXX
\end{definition}

Using backward induction, suppose that there were $k$ prior rejections, and the municipal resources shrank by $\delta^{k}R$. At this stage, the politicians are similar with regard to political affiliation, so that we drop the party label. A given councilor is recognized as the proposer with $1/N$ probability. The recognized councilor has to convince at least one-half of the council members to agree with her policy proposal. Therefore, a legislator $i$, that receives an offer from a proposer ($x_i, t_i$), at round $k$, will accept the offer if it is at least as good as waiting for the next round.

@@

We now have the entire strategy profile for the city councilors, provided that they reject the initial mayor's proposal. Proceeding by backward induction, we solve for the mayor's proposal that wins the support of the majority of the council. In the mayor's proposal policy concerns come into play and the offer for a government-aligned councilor is different than the offer for an opposition councilor. For a government aligned councilor, assuming the councilor is the pivotal for the implementation, the accepted offer $(x_m^g, t_m^g)$ has to be better than rejecting the proposal, and waiting for the next round:

@@

Note that $\underline{x}(1)$ is the offer of private goods if the councilor rejects the mayors offer, so that $k = 1$. As the benefit of retention $B$ is higher than the cost of providing the targeted goods $\underline{t}$, the optimal provision of targeted goods satisfy the threshold for reelection: $t^* = t_m^g = t_i = \underline{t}$. Rewriting the equations leads us to:

@@

And after some algebra:

\[
x_m^g \ \geq \ -p +\dfrac{1}{2N+1}\left[2\delta R \left(1 + \dfrac{1-\delta}{2N}\right)-N(\underline{t} + B) + B\right] \ \equiv \ \underline{x}_m^g
\]

For the opposition councilor, the offer has to compensate the political loss of having a mayor that is not aligned with the councilor's ideology. The optimal offer satisfy:

\[
-p + x_o^g + B \ \geq \ \dfrac{1}{N}\left( \delta R -\dfrac{N}{2}(\underline{x}(1) + \underline{t}) + B)  \right) + \left(1 - \dfrac{1}{N}\right)\left(\dfrac{1}{2}\right)(\underline{x}(1) + B)
\]

And after some algebra:

\[
x_o^g \ \geq \ p + \dfrac{1}{2N+1}\left[2\delta R \left(1 + \dfrac{1-\delta}{2N}\right)-N(\underline{t} + B) + B\right] \ \equiv \ \underline{x}_o^g
\]

\begin{lemma}
The offer for a government-aligned councilor is always lower than the offer for an opposition councilor.
\end{lemma}

\begin{proof}
Note that the $\underline{x}_m^g = 2p + \underline{x}_o^g$. As the political disagreement is higher than zero, $\underline{x}_m^g = - 2p + \underline{x}_o^g < \underline{x}_o^g$. Moreover, note that $\underline{x}_m^g - \underline{x}_o^g = -2p$.
\end{proof}

This characterizes the offer that the mayor will make to win the support of the council. Now, let the number of government politicians be equal to $\eta_g$, and the ex-ante probability of electing a councilor aligned with the government equal to $\gamma \in (0,1)$. We define the cost to govern $C_g(N)$ as XXXX:

\[
C_g(N) = \dfrac{N}{2}(\underline{t} + \underline{x}_o^g) + \sum_{\eta_g = 0}^{N/2}\mathbb{P}(\eta_g)\left( \underline{x}_m^g - \underline{x}_o^g \right)
\]

And the probability that $\eta_g$ councilors belong to the mayor's coalition is equal to $\mathbb{P}(\eta_g) = \binom{N}{\eta_g}\gamma^{\eta_g}(1-\gamma)^{N-\eta_g}$, which is the binomial probability given that the ex-ante chance to elect an aligned councilor is equal to $\gamma$.
